### 2011-02-10 paper/fitnessEvaluation.tex
edits

In order to measure the start and end position in the same pose, we
linearly interpolated the motion of the robot between the ready
position and the commanded gait, $g(t)$.  As shown in
\figref{linearInterp}, motion for the first second was interpolated
between the ready position and $g(t)$, for the next nine seconds the
robot moved exactly as specified by $g(t)$, and for the final two
seconds its motion was interpolated smoothly back to the ready
position.

\acmFig{linearInterp}{1}{To ensure fair fitness evaluations with as
  little noise as possible, the position of the robot was measured at
  the beginning and end of each run (red circles) in the same
  \emph{ready} pose (\figref{robot_close.jpg}).  Motion was
  interpolated linearly between this stationary pose and the commanded
  gait g(t) for one second at the beginning of the run and two seconds
  at the end, as shown above.}

>>>>

In order to measure the start and end position in the same pose, and to ensure fair fitness evaluations with as little noise as possible, we
linearly interpolated the motion of the robot between the ready
position and the commanded gait, $g(t)$.  As shown in
\figref{linearInterp}, motion for the first second was interpolated
between the ready position and $g(t)$, for the next nine seconds the
robot moved exactly as specified by $g(t)$, and for the final two
seconds its motion was interpolated smoothly back to the ready
position.

\acmFig{linearInterp}{1}{Linear interpolation from the ready position to the first and last commanded gait positions. The position of the robot was measured at
  the beginning and end of each run (red circles) in the same
 ready pose.  Motion was
  interpolated linearly between this stationary pose and the commanded
  gait g(t) for one second at the beginning of the run and two seconds
  at the end, as shown above.}

### 2011-02-10 paper/results.tex
edits

Before attempting to use the learning methods
(\secref{learningMethods}) to optimize the chosen family of
parameterized gaits (SineModel5, \secref{motionModel}), we did a brief
experiment to explore the five dimensions of the parameter space. To
accomplish this, we first selected a random parameter vector that resulted
in motion, but not an exceptional gait.

Then, in five different experiments, we varied each parameter
individually and plotted performance versus each changing parameter.
In addition, we duplicated each measurement to be able to estimate
standard deviation of the measurement noise at each point.  Results
for this exploration are shown in \figref{explore_dim_1} through
\figref{explore_dim_5}.  As is shown in the figures, some dimensions
are smoother than others, and some measurements are fairly noisy.

@@

\edit{maybe useful vvvvv}

To be as fair as possible when obtaining the results presented in
\secref{results}, we picked three random points --- $\vec{\theta}_A$,
$\vec{\theta}_B$, and $\vec{\theta}_C$ --- and started all seven
methods at these points...

Each of the parameter ??? methods was run on 3 different initial
parameter vectors, in order to fairly compare the algorithms. We allowed 
each run for the parameter ??? methods to
continue until the results plateaued (no improvement for one third of
the policies seen so far). Three runs of HyperNEAT were completed,
each with a different initial seed and run for 20 generations.

\edit{$\wedge\wedge\wedge\wedge\wedge\wedge\wedge\wedge$}

The maximizing strategies implemented include uniform random hill
climbing, Gaussian random hill climbing, policy gradient descent, and
Nelder-Mead simplex. As a group, these strategies produced results
similar to gaits randomly generated using the SineModel5 model. None
of these strategies consistently outperformed the others in the three
trials; A different strategy generated the best gait in each
trial. Uniform random hill climbing produced the best gait (11.37 body
lengths/minute) in the first trial, with the next best performing
strategy (Nelder-Mead simplex) resulting in a gait of 8.51 body
lengths/minute. The best gait in the second trial was actually
generated randomly (17.26 body-lengths/minute). Gaussian random hill
climbing, policy gradient descent, and Nelder-Mead simplex all
produced similar best gaits in the second trial, each around 14 body
lengths/minute. In the third trial, Nelder-Mead simplex and Gaussian
random hill climbing significantly outperformed the other maximizing
strategies, resulting in gaits of 14.83 and 13.40 body lengths/minute
respectively. Overall, Nelder-Mead simplex performed the best,
producing an average best gait of 12.32 body lengths/minute. Gaussian
random hill climbing followed, with an average best gait of 10.03 body
lengths/minute.

@@

\acmFig{vectorA}{1}{Results for runs beginning with vector A.
  Linear regression worked significantly better than the other
  algorithms, resulting in a gait (27.58 body lengths/minute) that
  walked over twice as fast as the next best gait -- uniform random
  hill climbing at 11.37 body lengths/minute -- and 5.3 times better
  than the previous hand-coded gait.}

\acmFig{vectorB}{1}{Results for run starting with vector B.  All
  algorithms performed similarly.  The random method actually got
  lucky this time and produced the best gait (17.26 body
  lengths/minute) and uniform random hill climbing produced the worst
  (9.44 body lengths/minute). The remaining algorithms each produced a
  gait around 13 body lengths/minute.}

\acmFig{vectorC}{1}{Results for runs starting with vector C.
  Simplex and Gaussian random hill climbing each produced a gait that
  substantially outperformed the other algorithms, with around 15 body
  lengths/minute and 13 body lengths/minute, respectively, whereas the
  other algorithms returned gaits of less than 5 body
  lengths/minute.}

Based on the three trials, as shown in \figref{std_error}, linear
regression worked the best, followed by simplex. Gaussian random hill
climbing performed about as well as random, but it is unclear whether
Gaussian would continue to improve and become much better than random
search if allowed to run for more iterations. Uniform random hill
climbing and policy gradient descent performed similarly to each other
on average, but differed greatly on individual runs.

\acmFigDouble{std_error}{.6}{Average results for each method starting with
  parameter vectors A, B, and C.  Error bars are plotted showing the
  standard error.  Linear regression outperformed other methods,
  followed by simplex. Gaussian random hill climbing performed about
  as well as random, but it is unclear whether Gaussian would continue
  to improve and become much better than random search if allowed to
  run for more iterations. Uniform random hill climbing and policy
  gradient descent performed similarly to each other on average, but
  differed greatly on individual runs.}

>>>>

Before optimizing the chosen family of
parameterized gaits (\secref{motionModel}) with learning methods, we performed an experiment to explore the five dimensions of the parameter space. Specifically, we first selected a random parameter vector that resulted
in motion, but not an exceptional gait. In five different experiments, we then varied each parameter
individually and measured performance.
Each measurement was repeated to get a rough estimate
of the measurement noise at each point.  The results
of this exploration (Figures \ref{fig:explore_dim_1} through
\ref{fig:explore_dim_5}) reveal that some dimensions
are smoother than others and that the measurement noise is variable and can be quite high.

@@

The maximizing strategies implemented include uniform random hill
climbing, Gaussian random hill climbing, policy gradient descent, and
Nelder-Mead simplex. As a group, these strategies produced results
similar to gaits randomly-generated with the SineModel5 model. None
of these strategies consistently outperformed the others in the three
trials; A different strategy generated the best gait in each
trial. Uniform random hill climbing produced the best gait (11.37 body
lengths/minute) in the first trial, with the next best performing
strategy (Nelder-Mead simplex) resulting in a gait of 8.51 body
lengths/minute. The best gait in the second trial was actually
generated randomly (17.26 body-lengths/minute). Gaussian random hill
climbing, policy gradient descent, and Nelder-Mead simplex all
produced similar best gaits in the second trial, each around 14 body
lengths/minute. In the third trial, Nelder-Mead simplex and Gaussian
random hill climbing significantly outperformed the other maximizing
strategies, resulting in gaits of 14.83 and 13.40 body lengths/minute
respectively. Overall, Nelder-Mead simplex performed the best,
producing an average best gait of 12.32 body lengths/minute. Gaussian
random hill climbing followed, with an average best gait of 10.03 body
lengths/minute.

@@

%\acmFig{vectorA}{1}{Results for runs beginning with vector A.
%  Linear regression worked significantly better than the other
%  algorithms, resulting in a gait (27.58 body lengths/minute) that
%  walked over twice as fast as the next best gait -- uniform random
%  hill climbing at 11.37 body lengths/minute -- and 5.3 times better
%  than the previous hand-coded gait.}
%
%\acmFig{vectorB}{1}{Results for run starting with vector B.  All
%  algorithms performed similarly.  The random method actually got
%  lucky this time and produced the best gait (17.26 body
%  lengths/minute) and uniform random hill climbing produced the worst
%  (9.44 body lengths/minute). The remaining algorithms each produced a
%  gait around 13 body lengths/minute.}
%
%\acmFig{vectorC}{1}{Results for runs starting with vector C.
%  Simplex and Gaussian random hill climbing each produced a gait that
%  substantially outperformed the other algorithms, with around 15 body
%  lengths/minute and 13 body lengths/minute, respectively, whereas the
%  other algorithms returned gaits of less than 5 body
%  lengths/minute.}

\acmFigDouble{std_error}{.6}{Average results ($\pm $ SEM) for each of the parametrized learning methods.  Linear regression performed the best,
  followed by Nelder-Mead simplex. Gaussian random hill
climbing performed about as well as random search, but may have surpassed random
with additional iterations. Uniform random hill
climbing and policy gradient descent performed similarly to each other
and underperformed random search.}
